\NeedsTeXFormat{LaTeX2e}
\documentclass[a4paper,11pt]{article}
\usepackage[utf8]{inputenc}

\usepackage{booktabs}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}

\usepackage{marvosym}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{etoolbox}

\usepackage{tikz}

\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;} 

\setlength{\textheight}{9in}
\setlength{\textwidth}{6in}
\setlength{\oddsidemargin}{.25in}
\setlength{\topmargin}{-.5in} 

\patchcmd{\thebibliography}{\section*{\refname}}
    {\begin{multicols}{2}[\section*{\refname}\singlespace\footnotesize]}{}{}
\patchcmd{\endthebibliography}{\endlist}{\endlist\end{multicols}}{}{}

\hyphenation{itself}

\title{{\small 02935 Introduction to applied statistics and R for PhD students: }\\[1em]Project description: Weather and travel demand}
\author{Niklas Christoffer Petersen}

\begin{document}


\singlespace
\maketitle

\onehalfspacing

\section{Background}\label{ch:background}

Observing and predicting the demand for bus travel is of major impact for designing and operating an efficient public transport system in any urban area. It is a common understanding, that there are several external factors that impact the travel demand. Examples include weather, events, etc. It is however uncertain how much each factor really contributes to fluctuation in travel demand. 

\section{Problem and Data}\label{ch:data}
This project aims to shed light upon the impact of specifically weather as an external factor for bus travel demand. The goal is to visualize and understand the data described in the following, and to show if any significant correlations between the datasets exists. Finally an simple model for travel demand prediction should be build in R.

\begin{itemize}
    \item The Danish national smart-card ticketing system, \emph{Rejsekort}, which is installed in every vehicle, and/or the camera-based \emph{automatic people counting} (APC) systems that are installed in a subset of vehicles. Both systems can be seen as a source for the actual travel demand (e.g. passengers boardings).
    \item Historic weather data from either \emph{OpenWeatherMap}\footnote{https://openweathermap.org/} or \emph{Weather Underground}\footnote{https://www.wunderground.com/}. Impacting variables include temperature, precipitation of rain/snow, wind speed and/or humidity.
\end{itemize}

For the sake of simplicity the analysis is expected to be limited to a single geographical area (e.g. Copenhagen) and bus line.

\begin{spacing}{1}
  \bibliographystyle{ieeetr}
  \bibliography{../references/library}
\end{spacing}

\end{document}
