\NeedsTeXFormat{LaTeX2e}
\documentclass[a4paper,11pt]{article}
\usepackage[utf8]{inputenc}

\usepackage[affil-it]{authblk}
\usepackage{natbib}
\usepackage[toc,nonumberlist,sort=def]{glossaries}

\usepackage{setspace}

\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epsfig}

\usepackage{marvosym}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{etoolbox}

\usepackage{todonotes}

\usepackage{tikz}

\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}

\setlength{\textheight}{9in}
\setlength{\textwidth}{6in}
\setlength{\oddsidemargin}{.25in}
\setlength{\topmargin}{-.5in} 

\setcitestyle{authoryear, open={(},close={)}}
%\patchcmd{\thebibliography}{\section*{\refname}}
%    {\begin{multicols}{2}[\section*{\refname}\singlespace\footnotesize]}{}{}
%\patchcmd{\endthebibliography}{\endlist}{\endlist\end{multicols}}{}{}

\hyphenation{itself}

\title{{\small 02935 Introduction to applied statistics and R for PhD students: }\\[1em]Statistical report: Weather and travel demand}

\author{Niklas Christoffer Petersen}
\affil{Transport Modelling, Department of Management Engineering \\ Technical University of Denmark, 2800 Kongens Lyngby, Denmark}

%\affil{Trafikselskabet Movia \\ Technical University of Denmark, 2800 Kongens Lyngby, Denmark}

\newglossaryentry{D_i}
{
  name=\ensuremath{D_i},
  description={is the \emph{observed} travel demand at the $i$'th observation/time, i.e.\ number of passenger boardings using the smart-card ticketing system}
}

\newglossaryentry{D_i_pred}
{
  name=\ensuremath{\widehat{D_i}},
  description={is the \emph{estimated} travel demand for time $i$}
}

\newglossaryentry{re_i}
{
  name=\ensuremath{\mathit{re}_i},
  description={is the \emph{relative error} of travel demand for time $i$}
}

\newglossaryentry{t_i}
{
  name=\ensuremath{t_i},
  description={is the timestamp for the $i$'th observation in local time (e.g.\ ``2016-10-01 00:00'')}
}

\newglossaryentry{dow_i}
{
  name=\ensuremath{\mathit{dow}_i},
  description={is the day of week of the $i$'th observation (i.e.\ \textsc{Mon}, \textsc{Tue}, ... \textsc{Sun})}
}

\newglossaryentry{hour_i}
{
  name=\ensuremath{\mathit{tod}_i},
  description={is the time of day  (i.e.\ 0 = ``00--01'', 1 = ``01--02'', ..., 23 = ``23--24'')}
}

\newglossaryentry{day_i}
{
  name=\ensuremath{\mathit{day}_i},
  description={is the day number of the $i$'th observation, i.e.\ number of days between $t_0$ and $t_i$}
}

\newglossaryentry{daytype_i}
{
  name=\ensuremath{\mathit{dt}_i},
  description={is the day type of the $i$'th observation, i.e.\ \textsc{Weekday} or \textsc{Weekend}}
}

\newglossaryentry{peek_i}
{
  name=\ensuremath{\mathit{peek}_i},
  description={is the peek class of the $i$'th observation, i.e.\ \textsc{No Peek}, \textsc{Morning}, or \textsc{Afternoon}. \textsc{Morning} peeks are \textsc{Weekdays} between 7--9, while \textsc{Afternoon} peeks are \textsc{Weekdays} between 15--18}
}

\newglossaryentry{temp_i}
{
  name=\ensuremath{\mathit{temp}_i},
  description={is the temperature in degree Celsius of the $i$'th observation}
}

\newglossaryentry{pptn_i}
{
  name=\ensuremath{\mathit{pptn}_i},
  description={is the precipitation for the entire day in mm of the $i$'th observation}
}

\newglossaryentry{ws_i}
{
  name=\ensuremath{\mathit{ws}_i},
  description={is the wind speed in kilometers per hour of the $i$'th observation}
}

\newglossaryentry{rh_i}
{
  name=\ensuremath{\mathit{rh}_i},
  description={is the relative humidity in percentage of the $i$'th observation}
}

\newglossaryentry{cond_i}
{
  name=\ensuremath{\mathit{cond}_i},
  description={is the weather condition of the $i$'th observation, i.e.\ \textsc{Clear}, \textsc{Cloudy}, \textsc{Overcast}, \textsc{Light Rain}, \textsc{Rain}, \textsc{Heavy Rain}, \textsc{Snow}, or \textsc{NA}}
}


\makeglossaries

\begin{document}
\singlespace
\maketitle
\thispagestyle{empty}
\clearpage

\onehalfspacing
\pagenumbering{arabic}
\tableofcontents
\clearpage
\glsaddall
\printglossaries


\clearpage

\section{Background}\label{ch:background}

Observing and predicting the demand for bus travel is of major impact for designing and operating an efficient public transport system in any urban area. It is a common understanding, that there are several external factors that impact the travel demand. Examples include weather, events, etc. It is however uncertain how much each factor really contributes to fluctuation in travel demand. 


\subsection{Related work}\label{ch:relatedWork}
The impact of weather conditions on transport is a well-studied area within the scientific field of transport modeling. Notable work include \citet{Lam2008}, who presents a model for link travel time using weather conditions with different rainfall intensities. Their work does not focus on public transportation system, but only on the road network level, and their modeling approach is a self-designed, domain specific, method for estimating OD (Origin/Destination) travel demand.
Likewise \citet{Yo2010} presents two-stage models for travel time prediction by composing respectively support vector machines (SVM) and artificial neural networks (ANN) with the Kalman filtering technique. They do include weather conditions, but only in the form of binary indicators for sunny days and rainy days are considered. 

\subsection{Objective and scope}\label{ch:objective}
This reports aims to shed light upon the impact of specifically weather as an external factor for bus travel demand. The goal is to visualize and understand the data described in \Cref{ch:desc}, and to show if any significant correlations between the datasets exists.

\begin{itemize}
    \item How much does different weather conditions and weather measurements impact the travel demand in the public transport system?
    \item Can increased/decreased demand be predicted using weather forecasts?
\end{itemize}

For the sake of simplicity the analysis is expected to be limited to a single geographical area (i.e. Copenhagen) and one bus line (i.e. 5A). Furthermore the period of analysis is restricted to October 1, 2016 to March 31, 2017. 

\subsection{Statistical programming and reproducibility}\label{ch:reproducibility}
The statistical programming for this work is implemented using the R programming language \citep{R}. All work conducted for this statistical report is open and freely available at \url{https://github.com/niklascp/02935-statistics-r}, including program source code and full result sets. This allows reproducibility of results and transparency of applied methods.
\clearpage

\section{Data description}\label{ch:desc}
As established earlier, the goal of this project is to shed light upon the impact of specifically weather as an external factor for bus travel demand. For this, both historical weather data, and measures of historical travel demand should be analyzed.

\input{desc_traveldemand.tex}

\input{desc_weather.tex}
\clearpage

\input{normalization.tex}
\clearpage

\input{statistical_analysis.tex}
\clearpage

\section{Discussion}
The use of a linear regression model (LR) for modeling the expected variations in the travel demand pattern, i.e.\ the normal condition, seem reasonable as supported by a low MAPE around 10 \%. It is though probably easy to improve this with more advanced regression models, e.g.\ support vector regression (SVR) or artificial neural networks (ANN), but this will compromise the interpretability of the simple LR-model.

The application of PCA did not yield any clear patterns of clusterings with respect to detection between increased and decreased demand deviation, and it is probably not appropriate method for this multivariate
problem, as the weather data set does not project well to any lower sub-spaces. This was shown by scree and loading plots. Instead an attempt for use of support vector machines (SVM) was made, and even though the model was not optimized (i.e. hyper-parameter tuning), it showed some capabilities for predicting high/low/normal demand from unseen weather data.

\subsection{Future work}
Future work should increase the precision of the prediction as well as test for changes between different geospatial areas (e.g.\ rural areas vs.\ suburban areas). Finally more external factors besides weather can be included, e.g.\ events.

\section{Conclusion}
This statistical report shed light upon the impact of specifically weather as an external factor for bus travel demand. The report has presented a model for measuring the deviations from the expected variations in the normal travel demand pattern.

Furthermore the correlation between this deviation measure, and weather has been analyzed using simple univariate analysis, unsupervised learning (PCA) and lastly supervised learning (SVM).

The overall result confirms a strong influence of weather conditions on the deviation measure, although is also implies that the impact is too complex for univariate analysis and requires supervised learning methods. It is suggested that further research should be invested in the supervised learning methods, as early results shows promising capabilities of predicting the deviation ahead of time.

\clearpage
\begin{spacing}{1}
  \bibliographystyle{apalike}
  \addcontentsline{toc}{section}{References}
  \bibliography{../references/library}
\end{spacing}

\clearpage
\appendix
\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}
\renewcommand{\thesubsection}{\Alph{subsection}}

\input{appx_travel_demand_data.tex}
\clearpage
\input{appx_weather_data.tex}
\clearpage
\input{appx_model_assumptions.tex}

\end{document}
